Optimization of String Rewriting Operations for 3D Fractal Generation with Genetic Algorithms

### 2018-06-03 Optimization-of-String-Rewriting-Operations-for-3D-Fractal-Generation-with-Genetic-Algorithms/paper.tex
Initial version of the paper was completed.

\begin{figure}[h]
  \centering
  \includegraphics[width=1.0\linewidth]{pic01}
  \caption{Fractal generated by Jack Hodkinson.}
\label{fig:pic01}
\end{figure}

@@

A finite 3D space is represented as three-dimensional array of voxels in the model. Each voxel is encoded with a single long integer value, which consists of the RGB information for the color of the voxel. In this model 32 different transition 3x3x3 matrices are selected, each corresponding to the one of the 32 colors. The steps for fractal generation are given in Algorithm \ref{alg:algo01}.

\begin{algorithm}
\caption{Fractal generation algorithm.}\label{alg:algo01}
\begin{algorithmic}[1]
\Procedure{Generate}{$Level,Voxel$}
\If{$Level \leqslant 0$}
  \State return\label{alg:step04}
\EndIf
\State $v\gets Split(Voxel)$\label{alg:step01}
\For{each $v$}
  \State Paint($v$)\label{alg:step02}
  \State Generate($Leve$-1, $v$)\label{alg:step03}
\EndFor
\EndProcedure
\end{algorithmic}
\end{algorithm}
\FloatBarrier

@@

The individuals in the genetic algorithm are presented as 32 transition matrices each with size 3x3x3. The order of the matrices shows which transition will be applied according to used colors order. The steps of the optimization are illustrated in Algorithm \ref{alg:algo02}.

\begin{algorithm}
\caption{Genetic algorithm.}\label{alg:algo02}
\begin{algorithmic}[1]
\Procedure{Optimization}{}
\State $p \gets Initialize()$\Comment{Initialize population.}
\While{$TimeLimit > 0$}
  \State Selection($p$)\Comment{Select parents.}\label{alg:step05}
  \State Crossover($p$)\Comment{Crossover parents.}
  \State Mutation($p$)\Comment{Mutate children.}
  \State Evaluation($p$)\Comment{Evaluate offspring.}
\EndWhile
\EndProcedure
\end{algorithmic}
\end{algorithm}
\FloatBarrier

Tournament selection with arity of two is proposed. Binary crossover is taken as the most symmetric crossover. From each parent randomly are taken 50\% of the elements. Random change of a color is used for the mutation operator. The evaluation is done by application of each 32 transition matrices with the certain level of recursion depth. Euclidean distance between generated fractal and target shape is used as fitness function. The initial fractal is a single cube painted with the color with the first index in the order of the colors. Genetic algorithms are global heuristics optimization and this means that it does not matter what kind of color the initial voxel will have.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=1.0\linewidth]{pic02}
  \caption{Voxelized sphere.}
\label{fig:pic02}
\end{figure}
\FloatBarrier

>>>>

\begin{figure}[h!]
  \centering
  \includegraphics[width=1.0\linewidth]{pic01}
  \caption{Fractal generated by Jack Hodkinson.}
\label{fig:pic01}
\end{figure}

@@

A finite 3D space is represented as three-dimensional array of voxels in the model. Each voxel is encoded with a single long integer value, which consists of the RGB information for the color of the voxel. In this model 32 different transition 3x3x3 matrices are selected, each corresponding to the one of the 32 colors. The steps for fractal generation are given in Algorithm \ref{algorithm01}.

\begin{algorithm}
\caption{Fractal generation algorithm.}\label{algorithm01}
\begin{algorithmic}[1]
\Procedure{Generate}{$Level,Voxel$}
\If{$Level \leqslant 0$}
  \State return\label{alg:step04}
\EndIf
\State $v\gets Split(Voxel)$\label{alg:step01}
\For{each $v$}
  \State Paint($v$)\label{alg:step02}
  \State Generate($Leve$-1, $v$)\label{alg:step03}
\EndFor
\EndProcedure
\end{algorithmic}
\end{algorithm}
\FloatBarrier

@@

The individuals in the genetic algorithm are presented as 32 transition matrices each with size 3x3x3. The order of the matrices shows which transition will be applied according to used colors order. The steps of the optimization are illustrated in Algorithm \ref{algorithm02}.

\begin{algorithm}
\caption{Genetic algorithm.}\label{algorithm02}
\begin{algorithmic}[1]
\Procedure{Optimization}{}
\State $p \gets Initialize()$\Comment{Initialize population.}
\While{$TimeLimit > 0$}
  \State Selection($p$)\Comment{Select parents.}
  \State Crossover($p$)\Comment{Crossover parents.}
  \State Mutation($p$)\Comment{Mutate children.}
  \State Evaluation($p$)\Comment{Evaluate offspring.}
\EndWhile
\EndProcedure
\end{algorithmic}
\end{algorithm}
\FloatBarrier

Tournament selection with arity of two is proposed. Uniform crossover is taken as the most symmetric crossover. From each parent randomly are taken 50\% of the elements. Random change of a color is used for the mutation operator. The evaluation is done by application of each 32 transition matrices with the certain level of recursion depth. Euclidean distance between generated fractal and target shape is used as fitness function. The initial fractal is a single cube painted with the color with the first index in the order of the colors. Genetic algorithms are global heuristics optimization and this means that it does not matter what kind of color the initial voxel will have.

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=1.0\linewidth]{pic02}
  \caption{Voxelized sphere.}
\label{fig:pic02}
\end{figure}
\FloatBarrier

The genetic algorithm is used with the following parameters:

\begin{table}[h!]
\centering
\label{table01}
\begin{tabular*}{\textwidth}{|c@{\extracolsep{\fill}}|c|}
\hline 
\textbf{Parameter} & \textbf{Value} \\
\hline
\hline
elitism rate & 0.10 \\
\hline
crossover rate & 0.90 \\
\hline
mutation rate & 0.01 \\
\hline
number of individuals & 37 \\
\hline
number of variables &  864 \\
\hline
optimization minutes & 30 \\
\hline
\end{tabular*}
\vspace{2 mm}
\caption{Genetic algorithm parameters.}
\end{table}
\FloatBarrier

\begin{figure}[h!]
  \centering
  \includegraphics[width=1.0\linewidth]{pic03}
  \caption{Steps taken in the optimization process.}
\label{fig:pic03}
\end{figure}
\FloatBarrier

As it is shown in Fig. \ref{fig:pic03} the optimization process converges much more slower in the 3D space than what was achieved by Hodkinson in the 2D space.
